---
title: "Reproducible Research Course Project 1"
output: html_document
---

```{r}
library(knitr)
library(dplyr)
library(ggplot2)
opts_chunk$set(echo = TRUE)
setwd('/Users/helstans/Documents')
```

## Loading and preprocessing the data

```{r}
# load data
data_row <- read.csv('activity.csv')

# remove NA in data
data <- data_row[ with (data_row, { !(is.na(steps)) } ), ]

# print header summary data
head(data)
```

## What is mean total number of steps taken per day?

```{r, echo=TRUE}
Daily_steps <- group_by(data, date)
Daily_steps2 <- summarise(Daily_steps, total = sum(steps))
Daily_steps2
```

```{r, hist, echo=TRUE}
hist(Daily_steps2$total, main="Number of steps per day", col = "blue")
```

```{r, echo=TRUE}
summary(Daily_steps2)
```

## What is the average daily activity pattern?
```{r, echo=TRUE}
stepsinterval <- aggregate(steps ~ interval, data, mean)
plot(stepsinterval$interval, stepsinterval$steps, type='l', 
     main="Average steps per day")
```

```{r, echo=TRUE}
max <- which.max(stepsinterval$steps)

# find interval with this max
stepsinterval[max, ]
```

## Imputing missing values

```{r, echo=TRUE}
sum(is.na(data_row))
```

Filling in all the missing values using the mean of that 5 minute interval
```{r, echo=TRUE}
data_imputed <- data_row
for (i in 1:nrow(data_imputed)) {
  if (is.na(data_imputed$steps[i])) {
    interval_value <- data_imputed$interval[i]
    steps_value <- stepsinterval[
      stepsinterval$interval == interval_value,]
    data_imputed$steps[i] <- steps_value$steps
  }
}
```

```{r, echo=TRUE}
df_imputed_steps_by_day <- aggregate(steps ~ date, data_imputed, sum)
head(df_imputed_steps_by_day)
```

```{r, echo=TRUE}
hist(df_imputed_steps_by_day$steps, main="Histogram of total number of steps per day with filled missing values")
```

```{r, echo=TRUE}
mean(df_imputed_steps_by_day$steps)
```

```{r, echo=TRUE}
median(df_imputed_steps_by_day$steps)
```

```{r, echo=TRUE}
mean(Daily_steps2$total)
```

```{r, echo=TRUE}
median(Daily_steps2$total)
```
means remain the same, but there is a difference in median value

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
data_imputed['type_of_day'] <- weekdays(as.Date(data_imputed$date))
data_imputed$type_of_day[data_imputed$type_of_day  %in% c('Saturday','Sunday') ] <- "weekend"
data_imputed$type_of_day[data_imputed$type_of_day != "weekend"] <- "weekday"
```

```{r, echo=TRUE}
# convert type_of_day
data_imputed$type_of_day <- as.factor(data_imputed$type_of_day)

# calculate average daily steps by interval
df_imputed_steps_by_interval <- aggregate(steps ~ interval + type_of_day, data_imputed, mean)

# plot
qplot(interval, 
      steps, 
      data = df_imputed_steps_by_interval, 
      type = 'l', 
      geom=c("line"),
      xlab = "Interval", 
      ylab = "Number of steps",
      col = "red",
      main = "Weekdays vs Weekends") +
  facet_wrap(~ type_of_day, ncol = 1)
```


